Concurrent Programming in Harmony

### 2021-08-08 book/paper.tex
Latest

\glsadd{state machine replication}%
\index{state machine replication}%
Besides availability, it is usually important that the replicated
services acts as if it were a single one.  This requires that the
replicas of the service coordinate their actions.
The \emph{Replicated State Machine Approach}~\cite{Lam78,S90} is a
general approach to do just this.
First, you model your service as a deterministic state
machine.  The replicas each run a copy of the state machine, started
in the same state.  As long as the replicas handle the same inputs
in the same order, determinism guarantees that they produce the same
outputs in the same order.
The trick then is to ensure that all replicas handle the
same requests in the same order and to do so in a way that continues
to work even if some subset of replicas crash.

\emph{Chain Replication}~\cite{vRS04} is such a replication protocol.
In Chain Replication, the replicas are organized in a linear chain.
Each replica monitors its direct predecessor in the chain.  If it has
no predecessors, we call it the \emph{head}.  The head may change over
time, as a head may crash and replaced by another.
We model the state machine as a \emph{history}: a sequence of operations.
Only the head is allowed to introduce new updates.
When it does so, it advertises the new update along with its position
in the history.  When the direct successor receives the next update
it is expecting, it appends the update to its history and likewise
advertises the new update.  When the update is applied to the history
of the \emph{tail} replica (the replica that has no live successors),
then the update is considered final.

So, when a replica fails, its successors should find out about it.
In practice, one server can detect the failure of another server by
pinging it.  If a server does not receive a response to its ping within
some maximum amount of time, then the server considers its peer crashed.
Note that this, in general is not a safe thing to do---the network or
the peer may be temporarily slow but the peer is not necessarily crashed
when the timer expires.
Nonetheless, we will assume here that failure detection does not make
mistakes and that eventually every failure is detected.
This is sometimes called the \emph{Fail-Stop} failure model~\cite{SS83},
which is distinct from the often more realistic \emph{Crash} failure model
where processes can crash but accurate detection is not available.
We will consider that failure model in the upcoming chapters.

\autoref{fig:chain} shows an implemenation of chain replication.
The code starts \texttt{NREPLICAS} \texttt{replica} threads numbered
0 through $\mathtt{NREPLICAS} - 1$.  At least one of the replicas 
must be immortal---the code selects one of them although none of the
replicas is required to fail during execution.
In practice, of course, a replica does not know whether it is immortal
or not---it should just assume so.  The immortality is only used for
modeling the system.

Each replica maintains its history and a set of its predecessors.
It then loops until either it fails or it has applied all
\texttt{NOPS} operations to its history.
$\mathbf{choose}(\{ \mathit{immortal}, \textbf{True} \})$ can only
evaluate to \textbf{False} in case $\textit{immortal}$ is false.
What it does within the loop depends on whether it is the head
(has no alive predecessors) or not.
Because failure detection is accurate, at most one replica can think
it is the head at any time.  Moreover, when it has
detected all its predecessors having failed, eventually some replica
thinks it is the head.

Messages have the format $(\mathit{source}, \mathit{payload})$,
where \textit{source} is the identifier of the replica that generated
the message and \textit{payload} has information about the state of
the replica.  The payload can take one of two forms.  The first is
\texttt{.crash}, indicating that the source of the message has
crashed.  The code models a replica crashing by that replica broadcasting
this message on the network.  The other form of payload is a tuple
$(\mathit{index}, \mathit{operation})$, which means that the source
placed the given operation at the given index in its history.

If a replica thinks it is the head, it adds an operation to its
history.
In practice this would be some operation that the head would have received
from a client.  Clients send their operations to the head, which adds them
to its history in the order received.
We are not modeling clients of the service.
Instead, in the model, we add \textit{self} to the history until the
history have \texttt{NOPS} operations on it or until the head crashes,
whichever comes first.  The head also broadcasts this update, which
is intended for its successor.  Because of failures, it may be that
more than one replica is a successor of the head at different times.

Otherwise, the replica awaits a message from its direct predecessor.
If the payload is \texttt{.crash}, then the replica removes this
predecessor from its set of predecessors.  For example, if there
are initially three replicas numbered 0, 1, and 2, and replica~1
crashes, then replica~2 removes~1 from its set of predecessors, making
replica~0 its direct (and only) predecessor.  If, instead, the payload
is a tuple, then it must be the case that it contains the next operation
to add to its history.  The replica adds the operation to its history
and broadcasts the update, again intended for its successor.

>>>>

\glsadd{state machine replication}%
\index{state machine replication}%
\index{determinism}%
Besides availability, it is usually important that the replicated
service acts as if it were a single one.  This requires that the
replicas of the service coordinate their actions.
The \emph{Replicated State Machine Approach}~\cite{Lam78,S90} is a
general approach to do just this.
First, you model your service as a deterministic state
machine.  The replicas each run a copy of the state machine, started
in the same state.  As long as the replicas handle the same inputs
in the same order, determinism guarantees that they produce the same
outputs in the same order.
The trick then is to ensure that all replicas handle the
same requests in the same order and to do so in a way that continues
to work even if some strict subset of replicas crash.

\emph{Chain Replication}~\cite{vRS04} is such a replication protocol.
In Chain Replication, the replicas are organized in a linear chain.
Each replica monitors its direct predecessor in the chain.  If a replica has
no predecessors, we call it the \emph{head}.  The head may change over
time, as a head may crash and replaced by another, but at any point in
time there is only one head
(and only one \emph{tail}, possibly the same replica).
We model the state machine as a \emph{history}: a sequence of operations.
Only the head is allowed to introduce new updates.
When it does so, it advertises the new update along with its position
in the history.  When the direct successor receives the next update
it is expecting, it appends the update to its history and likewise
advertises the new update.  When the update is applied to the history
of the tail replica (the replica that has no live successors),
then the update is considered final.

So, when a replica fails, its successors should find out about it.
In practice, one server can detect the failure of another server by
pinging it.  If a server does not receive a response to its ping within
some maximum amount of time, then the server considers its peer crashed.
Note that this, in general, is not a safe thing to do---the network or
the peer may be temporarily slow but the peer is not necessarily crashed
when the timer expires.
Nonetheless, we will assume here that failure detection does not make
mistakes and that eventually every failure is eventually detected.
This is called the \emph{Fail-Stop} failure model~\cite{SS83},
which is distinct from the often more realistic \emph{Crash} failure model
where processes can crash but accurate detection is not available.
We will consider that failure model in the upcoming chapters.

\autoref{fig:chain} shows an implemenation of chain replication.
The code starts \texttt{NREPLICAS} \texttt{replica} threads numbered
0 through $\mathtt{NREPLICAS} - 1$, with initially 0 being the head
and $\mathtt{NREPLICAS} - 1$ being the tail.  At least one of the replicas 
must be immortal---the code selects one of them although none of the
replicas is required to fail during execution.
Of course, a replica does not know whether it is immortal
or not in practice---it should just assume so.
The immortality of one of the replicas is only used for modeling the system.

Each replica maintains its history and a set of its predecessors.
It then loops until either it fails or it has applied all
\texttt{NOPS} operations to its history.
$\mathbf{choose}(\{ \mathit{immortal}, \textbf{True} \})$ can only
evaluate to \textbf{False} in case $\textit{immortal}$ is false; otherwise
it will always evaluate to \textbf{True}.
What it does within the loop depends on whether it is the head
(has no alive predecessors) or not.
Because failure detection is accurate, at most one replica can think
it is the head at any time (and, if so, it is in fact the head).
Moreover, when it has
detected all its predecessors having failed, eventually some replica
thinks it is the head.

Messages have the format $(\mathit{source}, \mathit{payload})$,
where \textit{source} is the identifier of the replica that generated
the message and \textit{payload} has information about the state of
the replica.  The payload can take one of two forms.  The first is
\texttt{.crash}, indicating that the source of the message has
crashed.  The code models a replica crashing by that replica broadcasting
a \texttt{.crash} message on the network.
The other form of payload is a tuple
$(\mathit{index}, \mathit{operation})$, which means that the source
placed the given operation at the given index in its history.

If a replica thinks it is the head, it adds an operation to its
history.
In practice this would be some operation that the head would have received
from a client.  Clients send their operations to the head, which adds them
to its history in the order received.
We are not modeling clients of the service.
Instead, in the model, the head adds its identifier \textit{self} to the
history until the
history has \texttt{NOPS} operations on it or until the head crashes,
whichever comes first.  The head also broadcasts each update, which
are intended for its successor.  Because of failures, it may be that
at different times a replica has different successors, hence the
broadcast.

Otherwise, the replica does not believe it is the head and
awaits a message from its direct predecessor.
If the payload is \texttt{.crash}, then the replica removes this
predecessor from its set of predecessors.  For example, if there
are initially three replicas numbered 0, 1, and 2, and replica~1
crashes, then replica~2 removes~1 from its set of predecessors, making
replica~0 its direct (and only) predecessor.  If, instead, the payload
is a tuple, then it must be the case that it contains the next operation
to add to its history.  The replica adds the operation to its history
and broadcasts the update, again intended for its successor.
